---
title: "Mini-Project 2"
author: "Dianne Caravela, Neyda Carrillo, Carmen Canedo"
date: "2019-03-24"
output: 
  html_document:
    code_folding: hide
---

  In the United States’ two-party system, many states lean towards one political party rather than the other. These states, referred to as “blue states” or “red states”, tend to vote for the Democratic Party or Republican Party, respectively. Although many states consistently vote along party lines, there are several “swing states” where elections could go to either party. Both parties have committees in most states that contribute money to campaigns during federal elections. It is important to understand the relationship between the number of committees in each state and the amount of money they contribute, as this could affect the outcome of federal elections.
  The current analysis examined the following questions: 1) was there a difference in number of committees and amount of contributions by party, and 2) was a greater number of committees associated with a greater number of contributions? The analysis used data collected by the Federal Election Commission during the 2011-2012 federal election cycle. The graph(s) below show the number of committees by state and the amount of contributions the committees gave to Democratic and Republican parties. From these data, we can learn how political parties differ in total number of committees and amount of contributions, as well as the relationship between these two. 
  In most states, one party was substantially higher than the other in committees and contributions. California, a typically blue state, had significantly more Democratic committees and contributions than the Republican party. With 967 committees and $1,528,437 in contributions, the Democratic committees gave $814,272 more than their Republican counterparts. Even in a swing state like Ohio, there were differences by party, with 187 committees and $157,795 for Democrats and 320 committees and $564,053 for Republicans. There were a few exceptions. Florida, another swing state, had similar amounts of committees and contributions for both parties, and five states only had committees for one party: Alaska, Idaho, Kansas, Montana, and Wyoming.
  There appeared to be an overall positive correlation between number of committees and amount of contributions. In 39 out of the 45 states that had data for both parties, parties with the greater number of committees also had the higher amount of contributions. For example, Massachusetts, a blue state, had 241 Democratic committees that contributed $353,073 and 8 Republican committees that contributed $7,650. Conversely, Texas, a traditionally red state, had 311 Democratic committees that contributed $468,767 and 848 Republican committees that contributed $1,673,242. The six states that did not follow this pattern were Florida, Hawaii, North Dakota, Tennessee, Utah, and West Virginia. For instance, Tennessee had 155 Democratic committees and 96 Republican committees, but the Republican committees contributed $58,078 more than the Democratic committees.
  These findings suggest that there are clear differences by party in the number of committees and amount of contributions. However, which party had the highest level depended on whether it was a red or blue state. As well, a strong pattern was observed that parties with more committees in a state had more contributions to campaigns. This leads to a possible question to examine for future research: which has a greater effect on the likelihood of winning federal elections, the number of committees or the contributions those committees give? Overall, the results of this analysis demonstrate the depths to which the two-party system is embedded in American elections. Ideally, there would be a balance in the amount of money going into each party’s campaigns, but based on these data, one party or the other tends to dominate in a state in terms of contributions.

Word count: 595

Link to GitHub^[https://github.com/ncarrillo1021/sds192-mp2]

```{r, include = FALSE, message=FALSE}
library(tidyverse)
library(devtools)
library(usmap)
library(gridExtra)
load("house_elections.rda")
load("candidates.rda")
load("committees.rda")
load("contributions.rda")
```


```{r, message=FALSE}
wrangle_state <- function(state_arg) {
  
#Creates list of democratic committees in given state
  party_cmte_DEM <- committees %>%
    filter(cmte_state == state_arg) %>% #filters for specific state
    filter(cmte_party_affiliation == "DEM") %>% #filter for democratic committees
    group_by(cmte_party_affiliation) %>%
    select(cmte_id)
  
#Creates list of republican committees in given state
  party_cmte_REP <- committees %>%
    filter(cmte_state == state_arg) %>%
    filter(cmte_party_affiliation == "REP") %>%
    group_by(cmte_party_affiliation) %>%
    select(cmte_id) 

#Creates table with just committee ID and transaction info for every transaction 
  trans_amt <- contributions %>%
    select(cmte_id, transaction_amt)

#Joins Rebublican committee list with transactions, summarizes the number of committies and the total amount of money they recieved  
  cmte_expenditure_DEM <- inner_join(party_cmte_DEM, trans_amt)  %>%
    summarize(num_committees = n(), total_cont = sum(transaction_amt))
  
#Does the same with Democratic committee list  
  cmte_expenditure_REP <- inner_join(party_cmte_REP, trans_amt)  %>%
    summarize(num_committees = n(), total_cont = sum(transaction_amt))

#Joins Republican and Democratic summary tables 
  final_table <- full_join(cmte_expenditure_DEM, cmte_expenditure_REP) %>%
    mutate(total_cont_thousands = total_cont / 1000)
 
#Creates graph of total contributions by party 
  total_cont_graph <- ggplot(final_table, aes(x = cmte_party_affiliation, y = total_cont_thousands, fill = cmte_party_affiliation)) +
    geom_bar( stat = "identity", width = 0.5) +
    scale_fill_manual(breaks = c("DEM", "REP"),
                    values = c("blue", "red")) + #sets colors
    guides(fill = FALSE) +
    xlab("Party Affiliation") + ylab("Total contributions (in thousands of dollars)") + ggtitle("Amount of contributions \nby committees by party in ", state_arg) #sets titles

#Creates graph of number of committees by party
  num_commit_graph <- ggplot(final_table, aes(x = cmte_party_affiliation, y = num_committees, fill = cmte_party_affiliation)) +
    geom_bar( stat = "identity", width = 0.5) +
    scale_fill_manual(breaks = c("DEM", "REP"),
                    values = c("blue", "red")) +
    guides(fill = FALSE) +
    xlab("Party Affiliation") + ylab("Number of Committees") + ggtitle("Number of committees \nby party afffiliation in ", state_arg)
 
#prints the two graphs side by side 
  grid.arrange(num_commit_graph, total_cont_graph, nrow = 1)
}


states <- c(states <- c("AL", "AK", "AZ", "AR", "CA", "CO", "CT", "DE", "FL", "GA", "HI", "ID", "IL", "IN", "IA", "KS", "KY", "LA", "ME", "MD", "MA", "MI", "MN", "MS", "MO", "MT", "NE", "NV", "NH", "NJ", "NM", "NY", "NC", "ND", "OH", "OK", "OR", "PA", "RI", "SC", "SD", "TN", "TX", "UT", "VT", "VA", "WA", "WV", "WI", "WY"))

sapply(states, wrangle_state) #applies the function to all 50 states

```

> Word count: 'r wordcountaddin::word_count()'


